---
title: "assignment1"
author: "Lucas Schiffer"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r include=FALSE}
library(readxl)
library(magrittr)
library(dplyr)
library(tidyr)
library(knitr)
```

## Introduction

This vignette is part of a large R package concerning assignment1 for BIOS 621; 
a GitHub repository containing the package can be found at 
https://github.com/schifferl/assignment1. Similarly, the build html vignette can
be found on RPubs at http://rpubs.com/schifferl/assignment1. Both links 
represent work in progress and should be interpreted accordingly. This vignette 
intends to reproduce the results given in the 2011 Chou et al. paper^[Chou, B., 
Krill, L. S., Horton, B. B., Barat, C. E. & Trimble, C. L. Disparities in Human 
Papillomavirus Vaccine Completion Among Vaccine Initiators. Obstetrics & 
Gynecology 118, 14–20 (2011).].

## Methods

$$\ln(OR)=\beta_0+AgeGroup\ x_1+InsuranceType\ x_2+LocationType\ x_3+
PracticeType\ x_4+Race\ x_5$$

## Results

```{r}
read_excel("../data-raw/gardasil.xls") %>%
summarise("11-17" = sum(AgeGroup==0), "18-24" = sum(AgeGroup==1), 
          "Public" = sum(MedAssist==1), "Private" = sum(MedAssist==0),
          "Suburban" = sum(LocationType==0), "Urban" = sum(LocationType==1),
          "Pediatrics" = sum(PracticeType==0), 
          "Family practice" = sum(PracticeType==1), 
          "Gynecology" = sum(PracticeType==2), "White" = sum(Race==0), 
          "Minority" = sum(Race==1|Race==2), 
          "African American" = sum(Race==1), "Hispanic" = sum(Race==2), 
          "Other" = sum(Race==3)) %>%
gather() %>% 
kable(., col.names = c("", "n"))
```


```{r}
read_excel("./data-raw/gardasil.xls") %>%
glm(Completed ~ AgeGroup + InsuranceType + LocationType + PracticeType + Race,
    family = binomial(), data = .) %>%
summary()
```

## Discussion

## References
