---
title: "assignment1"
author: "Lucas Schiffer"
date: "September 29, 2016"
output: html_document
vignette: >
  %\VignetteIndexEntry{assignment1}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
css: style.css
---

```{r include=FALSE}
library(magrittr)
library(readxl)
library(dplyr)
library(knitr)
library(assignment1)
```

## Introduction

This vignette is part of a larger R package concerning assignment1 for BIOS 621; 
a GitHub repository containing the package can be found at 
https://github.com/schifferl/assignment1. Similarly, the built html vignette can
be found on RPubs at http://rpubs.com/schifferl/assignment1. This vignette 
intends to reproduce the results given in the 2011 Chou et al. paper^[Chou, B., 
Krill, L. S., Horton, B. B., Barat, C. E. & Trimble, C. L. Disparities in Human 
Papillomavirus Vaccine Completion Among Vaccine Initiators. Obstetrics & 
Gynecology 118, 14–20 (2011).].

## Methods

Data related to the paper were provided by Levi Waldron, PhD in Microsoft Excel 
format (i.e. a .xls file), along with basic formatting code from which 
appropriate labels could be deduced. Labels were determined and serialized in 
the R environment as is shown in the subsequent code block.

```{r}
labels_AgeGroup <- c("18-26", "11-17")
labels_Race <- c("White", "African American", "Hispanic", "Other/Unknown")
labels_Shots <- c("Received 1 Shot", "Received 2 Shots", "Received 3 Shots")
labels_Completed <- c("Noncompleter", "Completer")
labels_InsuranceType <- c("Public", "Private payer", "Hospital–based",
                          "Military")
labels_MedAssist <- c("Public", "Private")
labels_Location <- c("Johns Hopkins Odenton", "Johns Hopkins White Marsh",
                     "Johns Hopkins Outpatient Center",
                     "Johns Hopkins Bayview Medical Offices")
labels_LocationType <- c("Urban", "Suburban")
labels_PracticeType <- c("Pediatrics", "Family practice", "Gynecology")
```

With the serialization of the variable labels, the Excel file was then read in 
using the readxl^[Hadley Wickham (2016). readxl: Read Excel Files. R package 
version 0.1.1. https://CRAN.R-project.org/package=readxl] package and all the 
variables, save for age, were interpreted to be categorical. The endomorphism of
the `data_frame` class was used in the manipulation of the data between readxl 
and dplyr^[Hadley Wickham and Romain Francois (2016). dplyr: A Grammar of Data 
Manipulation. R package version 0.5.0. https://CRAN.R-project.org/package=dplyr] 
commands. Syntax used in the serialization of data and interpolation of variable
labels is shown in the subsequent code block.

```{r}
gardasil <- read_excel("../inst/extdata/gardasil.xls") %>%
mutate(., AgeGroup = factor(AgeGroup, 1:0, labels_AgeGroup)) %>%
mutate(., Race = factor(Race, 0:3, labels_Race)) %>%
mutate(., Shots = factor(Shots, 1:3, labels_Shots)) %>%
mutate(., Completed = factor(Completed, 0:1, labels_Completed)) %>%
mutate(., InsuranceType = factor(InsuranceType, 0:3, labels_InsuranceType)) %>%
mutate(., MedAssist = factor(MedAssist, 1:0, labels_MedAssist)) %>%
mutate(., Location = factor(Location, 1:4, labels_Location)) %>%
mutate(., LocationType = factor(LocationType, 1:0, labels_LocationType)) %>%
mutate(., PracticeType = factor(PracticeType, 0:2, labels_PracticeType))
```

In order to reconstruct table 1 of the Chou et al. paper it was necessary to 
calculate sample standard deviation measures from the counts of categorical 
variables. The following formula^[https://stats.stackexchange.com/questions/
74797/calculating-standard-deviation-of-percentages] was used whereby $x$ 
represented a count within a category and $y$ represented the total of the 
category itself. Numerous functions were written so as to automate the process 
and a final reconstruction of table 1 was achieved using dplyr syntax.

$$s=\sqrt{\frac{1}{y} \frac{x}{y}\left(1-\frac{x}{y}\right)}$$

Similarly, the reconstruction of table 2 from the Chou et al. paper required a 
series of univariate generalized linear models from which the coefficients could 
be abstracted. The series of models was built as is subsequently specified and 
the reader may note that there is no variable $Minority$ previously devised in 
the serialization syntax. In order to produce results similar to the Chou et al. 
paper it was necessary to construct the $Minority$ variable from the $Race$ 
variable, considering $Minority$ to be either `"African American"` or 
`"Hispanic"`. In the case where $Race$ was `"Other/Unknown"` it was necessary to
exclude these cases from the analysis to obtain a result similar to Chou et al.

$$\ln(OR_{Completed})=\beta_0+AgeGroup\ x_1$$

$$\ln(OR_{Completed})=\beta_0+MedAssist\ x_1$$

$$\ln(OR_{Completed})=\beta_0+InsuranceType\ x_1$$

$$\ln(OR_{Completed})=\beta_0+LocationType\ x_1$$

$$\ln(OR_{Completed})=\beta_0+PracticeType\ x_1$$

$$\ln(OR_{Completed})=\beta_0+Minority\ x_1$$

$$\ln(OR_{Completed})=\beta_0+Race\ x_1$$

Finally, the reconstruction of table 3 required the construction of two 
multivariate logistic regression models - one that accounted for the missing 
cases within the $Minority$ category and another that excluded them. The results 
that are presented in the reconstructed table 3 represent coefficients from both 
models, given that it was necessary for reproduction. All but the $Minority$ 
coefficient come from the first model (i.e. where $\beta_6=Race$). There was a 
difference in number of observations between the two models by 186, with the 
model where $\beta_6=Race$ having the greater number. It did seem like a 
dishonest mistake, albeit perhaps committed honestly, and the implications are 
further discussed herein.

$$\ln(OR_{Completed})=\beta_0+AgeGroup\ x_1+MedAssist\ x_2+InsuranceType\ x_3+
LocationType\ x_4+PracticeType\ x_5+Race\ x_6$$

$$\ln(OR_{Completed})=\beta_0+AgeGroup\ x_1+MedAssist\ x_2+InsuranceType\ x_3+
LocationType\ x_4+PracticeType\ x_5+Minority\ x_6$$

It is noted that, in all models, the 95% confidence intervals were calculated 
using the standard error multiplied by the Z score for a probability density of 
0.95 taken to 7 significant figures. In the absence of further information in 
the Chou et al. paper as to how the confidence intervals were calculated, the 
method was conservatively selected.

## Results

To the exception of sample standard deviation percentages, it was possible to 
reproduce table 1 exactly as it was presented in the Chou et al. paper. There 
was a bit of experimentation necessary to determine that the 
`"Received 2 Shots"` category in fact contained anyone who had received 3 shots 
but had done so outside of the 12 month window and had thus been labeled a 
`"Noncompleter"`, in addition to those who had actually received 2 shots. The 
reconstructed table 1 is shown below, with sample standard deviation percentages
calculated as specified in the methods.

```{r echo=FALSE, message=FALSE, warning=FALSE}
Table_1() %>%
kable(col.names = c("", "n (%)", "Received 1 Shot", "Received 2 Shots", 
                            "Completer", "Noncompleter"),
      caption = "Table 1. Demographics of Study Cohort")
```

It was also possible to reproduce table 2 almost identically to the one 
presented in the Chou et al. paper, with slight divergences happening around the
confidence intervals and one of the p values. These differences are perhaps due 
to the exactness of the calculations and the internal implementations of the 
statistical methods within software used to calculate the confidence intervals. 
With the multiplication of the standard error by the Z score a confidence 
interval more narrow than than presented in the Chou et al. paper was achieved, 
as is shown in the reconstructed table 2 below.

```{r echo=FALSE, message=FALSE, warning=FALSE}
Table_2() %>%
kable(caption = "Table 2. Human Papillomavirus Vaccination Completion Rates, 
      Univariable Analysis (N=1,413)")
```

The reproduction of table 3 was unique among the others in that it was not 
possible to reproduce the results of Chou et al., even after significant 
experimentation. In the absence of sufficient detail for reproducibility, it was 
necessary to abstract coefficients from multiple models and results were 
approximately equal to those of Chou et al. but the exact results could not be 
attained. All coefficients and p values were within about 10% of the published 
values and are presented in the reproduced table 3 below.

```{r echo=FALSE, message=FALSE, warning=FALSE}
Table_3() %>% 
kable(caption = "Table 3. Multivariable Logistic Regression Analysis of 
      Predictors of Human Papillomavirus Vaccine Completion")
```

By the end of attempting to reproduce tables 1, 2, and 3 of the Chou et al. 
paper, it became clear that subtle differences persisted across all three tables
and no one table was an exact replica of the published variant. The differences 
are most likely attributable to use of statistical methods and software, with 
the reproduced and published results generally showing concordance.

## Discussion

As this and the Chou et al. analysis showed, $InsuranceType$ was found to be the
strongest predictor of Gardasil HPV vaccine series completion. Considering 
$InsuranceType$ a modifiable risk factor then, it begs the question of if 
modification would provide any shift in completion. To a limited extent perhaps 
it would, but it is rather more important to consider the situation whereby 
$InsuranceType$ is merely a proxy for socioeconomic status. In classical studies
of socioeconomic status, such as the Whitehall Study^[Marmot, M. G., Rose, G., 
Shipley, M. & Hamilton, P. J. Employment grade and coronary heart disease in 
British civil servants. J Epidemiol Community Health 32, 244–249 (1978).], it 
becomes clear that much of what dictates personal behavior is based in relation 
to feelings of autonomy. Furthermore, beyond the issues related to multiple 
collinearity^[Vatcheva, K. P., Lee, M., McCormick, J. B. & Rahbar, M. H. 
Multicollinearity in Regression Analyses Conducted in Epidemiologic Studies. 
Epidemiology (Sunnyvale) 6, (2016).], it would be important to consider 
$InsuranceType$ in the overall context of epidemiological studies and the 
multiplicity of complex exposures^[Patel, C. J. & Ioannidis, J. P. A. Placing 
epidemiological results in the context of multiplicity and typical correlations 
of exposures. J Epidemiol Community Health 68, 1096–1100 (2014).]. In the 
absence of further study that might validate the correlation of interest and 
validate its significance, it would be rash to assume causation on correlation 
alone.

## References
